---
title: "COVID-19 Cases in Toronto"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: https://github.com/jacobyokehongsi/starter_folder-main."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(knitr)
```

# Introduction

first para: motivational and broad
second para: what was done and what was found
3rd para: implications


The remainder of this paper is: Section \@ref(data) explains the data.

# Data

para or 2 introducing the dataset briefly

then show extract of dataset using kable (Table \@ref(tab:testing))

```{r testing}
covid_data <- 
  read_csv(here::here("inputs/data/covid_data.csv")) %>% 
  clean_names()

covid_data |>
  select(age_group, neighbourhood_name, episode_date) |>
  slice(1:10) |>
  kable(
    caption = "First ten rows of a dataset that shows age, neighbourhood and date",
    col.names = c("Age Group", "Neighbourhood", "Episode Date"),
    booktabs = TRUE, 
    linesep = ""
  )
```




Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)



Talk way more about it. 



# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


